---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%",
  fig.width = 16/2,
  fig.height = 9/2,
  message = FALSE,
  warning = FALSE,
  cache = TRUE
)
```

# forestecology 

[![Travis Build Status](https://travis-ci.org/rudeboybert/forestecology.svg?branch=master)](https://travis-ci.org/rudeboybert/forestecology)
[![Lifecycle: stable](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://www.tidyverse.org/lifecycle/#experimental)
[![CRAN status](https://www.r-pkg.org/badges/version/forestecology)](https://cran.r-project.org/package=forestecology)



## Installation

You can install the released version of forestecology from [CRAN](https://CRAN.R-project.org) with:

```{r, eval = FALSE}
install.packages("forestecology")
```

And the development version from [GitHub](https://github.com/) with:

```{r, eval = FALSE}
# install.packages("remotes")
remotes::install_github("rudeboybert/forestecology")
```



## Example analysis

We present an example analysis using:

* [Michigan Big Woods](https://doi.org/10.7302/wx55-kt18) research plot data
* [Smithsonian Conservation Biology Institute (SCBI) ForestGEO](https://github.com/SCBI-ForestGEO/SCBI-ForestGEO-Data/tree/master/tree_main_census) research plot data

```{r}
library(tidyverse)
library(forestecology)
library(snakecase)
library(skimr)
library(sf)
library(sfheaders)
# devtools::install_github("rvalavi/blockCV")
library(blockCV)
```



### Load & preprocess data

We first load and preprocess the data corresponding to the two census reflecting the time period over which we consider growth in dbh. We load them into R as "tibble" data frames thereby ensuring a standardized input/output format that can be used across all `tidyverse` packages @tidyverse. Furthermore, we ensure that the different variables have the correct names, types (`dbl`, `data`, `factor`). 

**Big Woods**:

```{r}
# Read in Big Woods census data from 2008 & 2014
bw_2008 <- 
  "https://deepblue.lib.umich.edu/data/downloads/z603qx485" %>% 
  read_delim(delim = "\t") %>% 
  mutate(spcode = to_any_case(spcode)) %>%
  select(
    treeID = treeid, stemID = stemtag, sp = spcode, quadrat, gx, gy, dbh, 
    date, codes
  )
bw_2014 <- 
  "https://deepblue.lib.umich.edu/data/downloads/1831ck00f" %>% 
  read_delim(delim = "\t") %>% 
  mutate(spcode = to_any_case(spcode)) %>%
  select(
    treeID = treeid, stemID = stemtag, sp = spcode, quadrat, gx, gy, dbh, 
    date, codes
  )

# Read in bigwoods grouping classification data
bw_species <- 
  "https://deepblue.lib.umich.edu/data/downloads/000000086" %>% 
  read_delim(delim = "\t") %>% 
  # convert all to snake case:
  mutate_at(c("species", "genus", "family", "idlevel", "spcode"), to_any_case) %>% 
  # join trait group
  left_join(families, by = c("spcode", "family")) %>% 
  mutate(
    sp = str_sub(genus, 1, 2), 
    sp = str_c(sp, str_sub(species, 1, 2)),
    sp = tolower(sp),
    latin = str_c(genus, species, sep = " "),
    latin = to_any_case(latin)
  ) %>% 
  select(sp = spcode, genus, species, latin, family, trait_group)

bw_2008 <- bw_2008 %>%
  left_join(bw_species,by='sp')

```

**SCBI**:

```{r}
scbi_2013 <- 
  "https://github.com/SCBI-ForestGEO/SCBI-ForestGEO-Data/raw/master/tree_main_census/data/census-csv-files/scbi.stem2.csv" %>% 
  read_csv() %>% 
  select(
    treeID, stemID, sp, quadrat, gx, gy, dbh, 
    date = ExactDate, codes, status
  )

scbi_2018 <-
  "https://github.com/SCBI-ForestGEO/SCBI-ForestGEO-Data/raw/master/tree_main_census/data/census-csv-files/scbi.stem3.csv" %>% 
  read_csv() %>% 
  select(
    treeID, stemID, sp, quadrat, gx, gy, dbh, 
    date = ExactDate, codes, status
  ) %>% 
  mutate(dbh = as.numeric(dbh))
```



### Compute growth

We then combine the two census data frames into a single data frame  that now includes a numerical variable `growth` reflecting the average annual growth of dbh in cm. Furthermore, variables that (in theory) remain unchanged between censuses appear only once, such as `gx`, `gy`, species-related variables. Variables that should change between censuses are tagged with `1/2` indicating earlier/later, such as `dbh1/dbh2`, `codes1/codes2`.

The resulting data frames are named with some variation of `growth_df`.

**Big Woods**:

```{r}
census_df1 <- bw_2008
# we need to filter out the resprouts
census_df2 <- bw_2014 %>% filter(!str_detect(codes,'R'))
id <- "treeID"

bw_growth_df <- 
  # Merge both censuses and compute growth:
  compute_growth(census_df1, census_df2, id) %>% 
  mutate(sp = to_any_case(sp)) %>% 
  left_join(bw_species, by = c("sp" = "spcode"))
```

**SCBI**:

```{r}
census_df1 <- scbi_2013
census_df2 <- scbi_2018
id <- "stemID"

scbi_growth_df <- 
  # Merge both censuses and compute growth:
  compute_growth(census_df1, census_df2, id) %>%
  # they are mesuaring in mm while we are measuring in cm!!!
  mutate(growth = growth/10)
```

```{r}
growth_df <- bind_rows(
  bw_growth_df %>% select(growth) %>% mutate(site = "bw"),
  scbi_growth_df %>% select(growth) %>% mutate(site = "scbi")
)
ggplot(growth_df, aes(x = growth, y = ..density.., fill = site)) +
  geom_histogram(alpha = 0.5, position = "identity", binwidth = 0.05) +
  labs(x = "Average annual growth in dbh (cm per yr)") +
  coord_cartesian(xlim = c(-0.5, 1))
```



### Add spatial information

We now add spatial information to the `growth_df` data frame. Specifically:

1. In order to control for study region "edge effects" (cite Waller?), we need a function that adds "buffers" of trees. In our case, since our model of interspecies competition relies on a spatial definition of who is a "neighboring competitor" and certain explanatory varibles such as biomass are cummulative, we need to ensure that all trees being modeled are not biased to have different neighbor structure, in particular the trees at the boundary of study regions.
1. Assign each tree to a "fold" for cross-validation purposes. Conventional cross-validation schemes assign observations to folds by resampling individual observations at random. However, underlying this scheme is an assumption that the observations are independent. In the case of forest census data, observations exhibit spatial autocorrelation, and thus this dependence must be incorporated in our resampling scheme @roberts2017 @pohjankukka2017. Packages that have implemented spatial cross-validation include @valavi2019

Thus, before we can incorporate the above information to `growth_df`, we need to define two constants:

```{r}
cv_fold_size <- 100
max_dist <- 7.5
```


#### Defining buffers

For a focal tree of interest, the `max_dist` of `r max_dist` meters acts as a radius defining a neighborhood within which all trees are considered competitors. In other words, all trees within `r max_dist` meters of the focal tree are considered competitor trees. Other studies have estimated this distance; we used `r max_dist` meters as an average of estimated values [@canham2004, @canham2006, @uriarte2004, @tatsumi2013]. Following Tobler's first law of geography that "everything is related to everything else, but near things are more related than distant things." @tobler1970firstlawofgeo, we assume that the degree of spatial autocorrelation is inversely-related to distance. However, we further assume that once trees are more than `r max_dist` meters apart, this autocorrelation is negligeable. 

**Example of buffer at work**:

```{r}
# Boundary polygon
square_boundary <- tibble(
  x = c(0,0,1,1),
  y = c(0,1,1,0)
) %>% 
  sf_polygon()

# Buffer polygon
square_buffer <- square_boundary %>% 
  st_buffer(dist = -0.1)

ggplot() +
  geom_sf(data = square_boundary) +
  geom_sf(data = square_buffer, col="red") 
```

**Big Woods**:

```{r}
# Bigwoods study region boundary polygon
bw_boundary <- bigwoods_study_region %>% 
  sf_polygon()

# Buffer polygon
bw_buffer <- bw_boundary %>%
  st_buffer(dist = -max_dist)

# Convert data frame to sf object
bw_growth_df <- bw_growth_df %>% 
  st_as_sf(coords = c("gx", "gy"))

# ID which points are in buffer and which are not
index <- st_intersects(bw_growth_df, bw_buffer, sparse = FALSE)
bw_growth_df <- bw_growth_df %>% 
  mutate(buffer = index)

# Plot
ggplot() +
  geom_sf(data = bw_boundary) +
  geom_sf(data = bw_buffer, col="red") +
  # Only random sample of 1000 trees:
  geom_sf(data = bw_growth_df %>% sample_n(1000), aes(col=buffer), size = 0.5)
```

**SCBI**:

```{r}
# Bigwoods study region boundary polygon
scbi_boundary <- tibble(x = c(0,400,400,0,0), y = c(0,0,640,640,0)) %>% 
  sf_polygon()

# Buffer polygon
scbi_buffer <- scbi_boundary %>%
  st_buffer(dist = -max_dist)

# Convert data frame to sf object
scbi_growth_df <- scbi_growth_df %>% 
  st_as_sf(coords = c("gx", "gy"))

# ID which points are in buffer and which are not
index <- st_intersects(scbi_growth_df, scbi_buffer, sparse = FALSE)
scbi_growth_df <- scbi_growth_df %>% 
  mutate(buffer = index)

# Plot
ggplot() +
  geom_sf(data = scbi_boundary) +
  geom_sf(data = scbi_buffer, col="red") +
  # Only random sample of 1000 trees:
  geom_sf(data = scbi_growth_df %>% sample_n(1000), aes(col=buffer), size = 0.5)
```


#### Spatial cross-validation

We use the [`blockCV`](https://github.com/rvalavi/blockCV) package to define the spatial grid, whose elements will act as the folds in our leave-one-out (by "one" we meen "one grid block") cross-validation scheme. 

**Big Woods**:

```{r}
bw_cv_grid <- spatialBlock(
  speciesData = bw_growth_df, theRange = 100, k = 28, 
  xOffset = 0.5, yOffset = 0.99999,
  verbose = FALSE
)

# Add foldID to data
bw_growth_df <- bw_growth_df %>% 
  mutate(
    foldID = bw_cv_grid$foldID,
    foldID = factor(foldID)
  )

# Visualize grid
bw_cv_grid$plots +
  geom_sf(data = bw_growth_df, aes(col=foldID), size = 0.1)
```

**SCBI**:

```{r}
scbi_cv_grid <- spatialBlock(
  speciesData = scbi_growth_df, theRange = 100, k = 28, yOffset = 0.6, verbose = FALSE
)

# Add foldID to data
scbi_growth_df <- scbi_growth_df %>% 
  mutate(
    foldID = scbi_cv_grid$foldID,
    foldID = factor(foldID)
  )

# Visualize grid
scbi_cv_grid$plots +
  geom_sf(data = scbi_growth_df, aes(col=foldID), size = 0.1)
```



### Model specification

Next we specify the growth model we want. We will model growth as a function of tree identiy, size, and the size and identiy of its neighbors. The model can be run on different grouping of individuals (e.g., based on species or trait groupings). It can also be run with different measures of competition. 

* `model_number = 1`: No competition growth only depends on `dbh` and individual grouping.
* `model_number = 2`: Competition but identity of competitor does not matter, just sum of competitior biomass.
* `model_number = 3`: Full model with competition and competitor identity.

```{r}
bw_specs <- get_model_specs(bw_2008, 3, 'trait_group')
bw_specs

scbi_specs <- get_model_specs(scbi_2013, 3, 'sp')

```

* `get_model_specs()`. Inputs:
    + define identification/grouping classification
    + `growth_df` format
* `create_focal_and_comp()`. Inputs:
    + `growth_df`
    + `max_dist`: who are your competitors
    + `model_specs`: grouping of competitor biomasses



### Model fit and prediction

* `fit_bayesian_model()`. 
    + Inputs: `focal_and_comp`, `model_specs`
    + Output: `posterior_param`
* `predict_bayesian_model()`. 
    + Inputs: `focal_and_comp`, `model_specs`, `posterior_param`
    + Output: `focal_and_comp` with updated `growth_hat` variable
* Show observed vs predicted growth
* CV: `run_cv()` is a wrapper to `fit_bayesian_model()` and `predict_bayesian_model()`. 
* Permutation test:
  + Add line to that shuffles competitor species within `focal_and_comp` just before fitting model
  + permutation AND CV: add `run_shuffle` argument to `run_cv()`



### Model output and performance

* Make plot posterior parameter functions





